---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-")
```

# NNDM

<!-- badges: start -->
<!-- badges: end -->

**Note:** An updated version of NNDM with enhanced capabilities has been included in the `CAST` package (available on CRAN).

The goal of the `NNDM` package is to provide tools to perform Nearest Neighbour Distance Matching (NNDM) Leave-One-Out (LOO) Cross-Validation (CV) to estimate map accuracy for a given prediction task. In NNDM LOO CV, the nearest neighbour distance distribution function between the test and training data during the LOO CV process is matched to the nearest neighbour distance distribution function between the target prediction and training points, for those distances in which autocorrelation is present. NNDM CV can be used for various prediction areas (geographic interpolation/extrapolation), sampling patterns (regular/random/clustered), and landscape autocorrelation ranges. The package also includes tools to compute buffered-LOO (bLOO) CV.


Reference:

C. Milà, J. Mateu, E. Pebesma, and H. Meyer, 'Nearest Neighbour Distance Matching Leave-One-Out Cross-Validation for map validation', *Methods in Ecology and Evolution* (2022).

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
library("devtools")
devtools::install_github("carlesmila/NNDM")
```

## Example 

In this example we will use LOO, bLOO and NNDM LOO CV for map validation in a spatial interpolation task  to illustrate the typical use of the package. We will use the `ranger` implementation of the Random Forest (RF) algorithm wrapped in the `caret` package, but the CV indices generated by `NNDM` can also be used with other machine learning packages.


### The meuse dataset

We will use the `meuse` dataset included in the `sp` package, which includes samples of topsoil zinc concentrations in a flood plain of the river Meuse, in the Netherlands. More information on the dataset can be obtained by using `?meuse`. 

```{r meuse dataset data, warning=FALSE, message=F}
library("NNDM")
library("caret")
library("sp")
library("sf")
library("knitr")
library("gstat")
library("gridExtra")

# Sample data
data("meuse")
meuse <- st_as_sf(meuse, coords = c("x", "y"), crs = 28992, remove = F)

# AOI polygon
data("meuse.area")
meuse.area <- SpatialPolygons(list(Polygons(list(Polygon(meuse.area)), "area")))
meuse.area <- st_as_sf(meuse.area)
st_crs(meuse.area) <- 28992

# Prediction grid
data("meuse.grid")
meuse.grid <- st_as_sf(meuse.grid, coords = c("x", "y"), crs = 28992, remove = F)
```

```{r samples-and-outcome, echo=FALSE, fig.width=8, fig.height=5, fig.align = "center"}
# Samples and outcome
p1 <- ggplot() +
  geom_sf(data = meuse.area, fill = "grey90", alpha = 0.25) +
  geom_sf(data = meuse, aes(col=zinc)) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  scale_colour_viridis_c() +
  ggtitle("AOI, samples, \nand measured outcome")

# Covariate 1: distance to the Meuse river
p2 <- ggplot() +
  geom_sf(data = meuse.grid, aes(col=dist), size = 0.5) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  scale_colour_viridis_c()  +
  ggtitle("Prediction grid and one of the covariates\n(distance to Meuse river)")

grid.arrange(p1, p2, nrow=1)
```

### LOO CV

We start by fitting a RF model with the selected predictors (no hyperparameter tuning is done in this example) and validating it with LOO CV, which is readily available in `caret`. 

```{r LOOCV, warning=F}
# Fit model
trainControl_LOO <- trainControl(method = "LOOCV", savePredictions = T)
paramGrid <-  data.frame(mtry = 2, min.node.size = 5, splitrule = "variance")
mod_LOO <- train(zinc ~ x + y + dist + ffreq + soil,
                 method = "ranger",
                 trControl = trainControl_LOO,
                 tuneGrid = paramGrid, 
                 data = meuse, 
                 seed=12345)
```

### Estimating autocorrelation ranges

Both bLOO and NNDM LOO CV require an autocorrelation range parameter to be estimated. There are many proposals on how to estimate the radius of bLOO/NNDM LOO CV, being the most frequent the residual or outcome autocorrelation ranges (see our article for more details). Here we estimate the residual autocorrelation range:

```{r rescor, warning=F, fig.width=5, fig.height=4, fig.align = "center"}
# Estimate variogram on the residual and return range
meuse$res <- meuse$zinc - predict(mod_LOO, newdata=meuse)
empvar <- variogram(res~1, data = meuse)
fitvar <- fit.variogram(empvar, vgm(model="Sph", nugget = T))
plot(empvar, fitvar, cutoff=1500, main = "Residual semi-variogram estimation")
(resrange <- fitvar$range[2]) # Residual autocorrelation range in m
```

### bLOO CV

With the estimated range, we can now compute the bLOO CV indices using the `bLOO` function. We print the object and observe that with that radius, the average training size of each bLOO iteration is of roughly 110 points. We can also plot the object for a given observation and explore which observations are used for training, testing, or are excluded. We use these indices in a `caret` model to perform a bLOO CV.

```{r bLOOCV, fig.width=4, fig.height=4, warning=F, fig.align = "center"}
# Compute bLOO indices
(bLOO_indices <- bLOO(meuse, resrange, min_train = 0.5))
# Plot for one CV iteration
plot(bLOO_indices, 53) +
  theme(axis.text = element_blank(), axis.ticks = element_blank())

# Evaluate RF model using bLOO CV
trainControl_bLOO <- trainControl(method = "cv", savePredictions = T,
                                  index=bLOO_indices$indx_train,
                                  indexOut=bLOO_indices$indx_test)
mod_bLOO <- train(zinc ~ x + y + dist + ffreq + soil,
                  method = "ranger",
                  trControl = trainControl_bLOO,
                  tuneGrid = paramGrid, 
                  data = meuse, 
                  seed=12345)
```

### NNDM LOO CV

Similarly, we compute NNDM LOO CV indices using the function `nndm`. When plotting the object, the estimated $G$ functions are shown. In this case, the nearest neighbour distance distribution function found during LOO CV $\hat{G}_j(r)$ is smaller than the nearest neighbour distance distribution found during prediction $\hat{G}_{i \, j}(r)$ for short distances, and is slightly larger for long distances (see our related paper for definitions of $G$ functions and more detail on their interpretation). $\hat{G}_{j}^{*} (r, \mathbf{L})$, i.e. the nearest neighbour distance distribution during NNDM LOO CV, overlaps with  $\hat{G}_j(r)$ for short distances but matches $\hat{G}_{i \, j}(r)$ for long distances. In any case, by printing the object we see that, in this case, the difference between LOO CV and NNDM CV is minimal, as on average 153.9 points out of 154 are used for training in each NNDM LOO CV iteration.

```{r NNDM, fig.width=5, fig.height=3, warning=F, fig.align = "center"}
# Compute NNDM indices
(NNDM_indices <- nndm(meuse, meuse.grid, resrange, min_train = 0.5))
# Plot NNDM functions
plot(NNDM_indices)

# Evaluate RF model using NDM CV
trainControl_NNDM <- trainControl(method = "cv", savePredictions = T,
                                  index=NNDM_indices$indx_train,
                                  indexOut=NNDM_indices$indx_test)
mod_NNDM <- train(zinc ~ x + y + dist + ffreq + soil,
                  method = "ranger",
                  trControl = trainControl_NNDM,
                  tuneGrid = paramGrid, 
                  data = meuse, 
                  seed=12345)
```

### Comparing results

We compute the different CV results by manually computing the statistics for bLOO and NNDM (`caret`results for LOO customised folds are not appropriate):

```{r compare results}
# Compute and return results
stats_LOO <- data.frame(validation="LOO CV",
                        RMSE=mod_LOO$results$RMSE, 
                        R2=mod_LOO$results$Rsquared)
stats_bLOO <- data.frame(validation="bLOO CV",
                         RMSE=with(mod_bLOO$pred, sqrt(mean((obs-pred)^2))), 
                         R2=with(mod_bLOO$pred, cor(obs, pred)^2))
stats_NNDM <- data.frame(validation="NNDM LOO CV",
                         RMSE=with(mod_NNDM$pred, sqrt(mean((obs-pred)^2))), 
                         R2=with(mod_NNDM$pred, cor(obs, pred)^2))
```

```{r compare results2, echo=F}
kable(rbind(stats_LOO, stats_bLOO, stats_NNDM), digits = 2)
```

In this example, we see that LOO and NNDM LOO CV return very similar results while bLOO CV results in a much lower estimated map accuracy. 
